---
title: "Event-based surveillance - network analysis (AI)"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=F}
suppressMessages(library(readxl))
library(reshape2)
library(dplyr)
library(ggplot2)
library(igraph)
library(stringr)
library(ggpubr)
library(readr)
library(RColorBrewer)
source("./sna_functions.R")
```


# 1. Data Import


```{r, message=FALSE, warning=F}

# Reports from PADI-web
REPORTS_PW <- read_delim("./dataverse/padiweb_reports.csv", show_col_types = FALSE) %>% mutate(publication_date = as.Date(publication_date, format="%Y-%m-%d"))

# Reports from HealthMap
REPORTS_HM <- read_delim("./dataverse/healthmap_reports.csv", show_col_types = FALSE) %>% mutate(publication_date = as.Date(publication_date, format="%Y-%m-%d"))

# Events empres-i database
OIE_EVENTS  <- open_official_AI("./dataverse/AI_official.csv", date_format =  "%Y-%m-%d")
OIE_EVENTS <- filter(OIE_EVENTS, reportingdate >= "2018-07-01") %>% filter(reportingdate <= "2019-06-30")

# Manually recorded non-official events
UNOFFICIAL_EVENTS <- read_delim("./dataverse/unofficial_events.csv", 
                        delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)

# Sources
SOURCES <- read_delim(file="./dataverse/sources.csv", delim=";", show_col_types = F) %>% mutate(label = str_trim(label))

# Tables for PADI-web network
EVENTS_PW <- read_delim("./dataverse/padiweb_events.csv", 
                        delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)
PATHS_PW <- read_delim("./dataverse/padiweb_paths.csv", delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)
PATHS_PW$source_1 <- str_trim(PATHS_PW$source_1)
PATHS_PW$source_2 <- str_trim(PATHS_PW$source_2)
PATHS_EVENTS_PW <- read_delim("./dataverse/padiweb_paths_events.csv", delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)

# Tables for HealthMap network
EVENTS_HM <- read_delim("./dataverse/healthmap_events.csv", 
                        delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)
PATHS_HM <- read_delim("./dataverse/healthmap_paths.csv", delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)
PATHS_HM$source_1 <- str_trim(PATHS_HM$source_1)
PATHS_HM$source_2 <- str_trim(PATHS_HM$source_2)
PATHS_HM$id_from_source_1 <- as.character(PATHS_HM$id_from_source_1)
PATHS_EVENTS_HM <- read_delim("./dataverse/healthmap_paths_events.csv", delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)

```


# 2. Descriptive statistics

## 2.1. Events

### Official events from Empres-i
```{r, message=FALSE, warning=F}
rbind(OIE_EVENTS %>% count(type_host) %>% 
  mutate(prop = prop.table(n)),
  c("total", nrow(OIE_EVENTS), 1))
```


### Events disseminated through the networks
```{r}
# PADI-web
EVENTS_PW <- EVENTS_PW %>% filter(id_event %in% PATHS_EVENTS_PW$id_event) 
rbind(EVENTS_PW %>% select(id_event, type_event, host) %>% distinct() %>% count(type_event) %>% 
    mutate(prop = round(prop.table(n),3)),
  c("total", nrow(EVENTS_PW), 1))

# HealthMap
EVENTS_HM <- EVENTS_HM %>% filter(id_event %in% PATHS_EVENTS_HM$id_event) 
rbind(EVENTS_HM %>% select(id_event, type_event, host) %>% distinct() %>% count(type_event) %>% 
  mutate(prop = round(prop.table(n),3)),
  c("total", nrow(EVENTS_HM), 1))
```

### Overlap of detected events
```{r}
# PADI-web
EVENTS_PW %>% filter(!id_event %in% EVENTS_HM$id_event) %>% select(id_event, type_event, type_host) %>% group_by(type_event) %>% distinct() %>% count(type_host)

# HealthMap
EVENTS_HM %>% filter(!id_event %in% EVENTS_PW$id_event) %>% select(id_event, type_event, type_host) %>% group_by(type_event) %>% distinct() %>% count(type_host)

```

## 2.2 Host

```{r}
# PADI-web
EVENTS_PW %>% select(id_event, type_event, type_host) %>% group_by(type_event) %>% distinct() %>% count(type_host)%>% 
  mutate(prop = round(prop.table(n),3))

# HealthMap
EVENTS_HM %>% select(id_event, type_event, type_host) %>% group_by(type_event) %>% distinct() %>% count(type_host)%>% 
  mutate(prop = round(prop.table(n),3))
```

## 2.3 Reports

### Types of reports
```{r, message=FALSE}
# PADI-web
rbind(REPORTS_PW %>% count(type) %>% 
  mutate(prop = prop.table(n)) %>% arrange(desc(prop)),
  c("total", nrow(REPORTS_PW), 1))

# HealthMap
rbind(REPORTS_HM %>% count(type) %>% 
  mutate(prop = prop.table(n)) %>% arrange(desc(prop)),
  c("total", nrow(REPORTS_HM), 1))
```


### Events of reports
```{r}
PATHS_EVENTS_PW  <- merge(PATHS_EVENTS_PW, EVENTS_PW)
REPORTS_EVENTS_PW <- merge(PATHS_EVENTS_PW, select(PATHS_PW, id_path, id_from_source_1))  %>% distinct()

event_by_report_pw <- aggregate(select(REPORTS_EVENTS_PW, id_from_source_1),
                                  by = list(id_article= REPORTS_EVENTS_PW$id_from_source_1), 
                                  FUN = length)


PATHS_EVENTS_HM  <- merge(PATHS_EVENTS_HM, EVENTS_HM)
REPORTS_EVENTS_HM <-merge(PATHS_EVENTS_HM, select(PATHS_HM, id_path, id_from_source_1))%>% distinct()
event_by_report_hm <- aggregate(select(REPORTS_EVENTS_HM, id_from_source_1),
                                  by = list(id_article = REPORTS_EVENTS_HM$id_from_source_1), 
                                  FUN = length)

summary(event_by_report_pw$id_from_source_1)
summary(event_by_report_hm$id_from_source_1)

```

## 2.4. Timeliness


```{r}
# Official events
id_events_off <- pull(filter(EVENTS_PW, type_event == "official"), id_event)

# Paths transmitting official events
id_paths_off <- unique(pull(filter(PATHS_EVENTS_PW, id_event %in% id_events_off), id_path))

PATHS_EVENTS_OFF_PW <- filter(PATHS_EVENTS_PW, id_path %in% id_paths_off) %>% filter(id_event %in% id_events_off)
  
PATHS_EVENTS_OFF_PW <- merge(PATHS_EVENTS_OFF_PW, EVENTS_PW, by.all = "id_event", all.y =F) 

PATHS_EVENTS_OFF_PW <- merge(PATHS_EVENTS_OFF_PW, select(PATHS_PW, id_path, source_1, date_edge), by.all = "id_path")

PATHS_EVENTS_OFF_PW <- PATHS_EVENTS_OFF_PW %>% filter(source_1 %in% c("PADI-web"))

PATHS_EVENTS_OFF_PW$timeliness <- as.numeric(difftime(PATHS_EVENTS_OFF_PW$date_edge, 
                                                    PATHS_EVENTS_OFF_PW$reportingdate,
     units="days"))


# For each event, detect the first report which detected it
df_timeliness_pw <- PATHS_EVENTS_OFF_PW %>% select(id_event, type_host, timeliness) %>%
  group_by(id_event, type_host) %>% 
  filter(timeliness == min(timeliness)) %>% unique()

summary(df_timeliness_pw$timeliness)

# Early detection
length(which(df_timeliness_pw$timeliness<0))

# Late detection
length(which(df_timeliness_pw$timeliness>=0))

```
```{r}
host_levels <-  c("domestic", "wild", "environmental", "unspecified")
df_timeliness_pw <- df_timeliness_pw %>% mutate(Host = factor(type_host, levels = host_levels)) %>% group_by(timeliness, Host) %>% count() %>% mutate(n = n*100/nrow(df_timeliness_pw))

nrow(filter(PATHS_EVENTS_OFF_PW, type_host == "wild") %>% filter(timeliness<0))
nrow(filter(PATHS_EVENTS_OFF_PW, type_host == "domestic") %>% filter(timeliness<0))
```

```{r}
# Official events
id_events_off <- pull(filter(EVENTS_HM, type_event == "official"), id_event)

# Paths transmitting official events
id_paths_off <- unique(pull(filter(PATHS_EVENTS_HM, id_event %in% id_events_off), id_path))

PATHS_EVENTS_OFF_HM <- filter(PATHS_EVENTS_HM, id_path %in% id_paths_off) %>% filter(id_event %in% id_events_off)
  
PATHS_EVENTS_OFF_HM <- merge(PATHS_EVENTS_OFF_HM, EVENTS_HM, by.all = "id_event", all.y =F) 

PATHS_EVENTS_OFF_HM <- merge(PATHS_EVENTS_OFF_HM, select(PATHS_HM, id_path, source_1, date_edge), by.all = "id_path")

PATHS_EVENTS_OFF_HM <- PATHS_EVENTS_OFF_HM %>% filter(source_1 %in% c("HealthMap"))

PATHS_EVENTS_OFF_HM$timeliness <- as.numeric(difftime(PATHS_EVENTS_OFF_HM$date_edge, 
                                                    PATHS_EVENTS_OFF_HM$reportingdate,
     units="days"))

# For each event, detect the first report which detected it
df_timeliness_hm <- PATHS_EVENTS_OFF_HM %>% select(id_event, type_host, timeliness) %>%
  group_by(id_event, type_host) %>% 
  filter(timeliness == min(timeliness)) %>% unique()

df_timeliness_hm <- df_timeliness_hm %>% mutate(Host = factor(type_host, levels = host_levels))


df_timeliness_hm <- df_timeliness_hm %>% select(-id_event) %>% group_by(timeliness, Host) %>% count() %>% mutate(n = n*100/nrow(df_timeliness_hm))


summary(PATHS_EVENTS_OFF_HM$timeliness)

# Early detection
length(which(PATHS_EVENTS_OFF_HM$timeliness<0))

# Late detection
length(which(PATHS_EVENTS_OFF_HM$timeliness>=0))
```
```{r}
nrow(filter(PATHS_EVENTS_OFF_HM, type_host == "wild") %>% filter(timeliness<0))
nrow(filter(PATHS_EVENTS_OFF_HM, type_host == "domestic") %>% filter(timeliness<0))
```

```{r, warning=F}
timeliness_pw <- create_timeliness_plot(df_timeliness_pw, min_timeliness = -35, max_timeliness = 30, size_font = 14)
timeliness_pw %>% ggexport(filename = "./figures/pw_timeliness.png", width = 800, height = 480)
timeliness_pw
```

```{r , warning=FALSE}
timeliness_hm <-create_timeliness_plot(df_timeliness_hm, min_timeliness = -35, max_timeliness = 30, size_font = 14)
timeliness_hm %>% ggexport(filename = "./figures/hm_timeliness.png", width = 800, height = 480)
timeliness_hm
```

## 2.5. Paths metrics

### Reactivity                        

```{r , echo=FALSE}
# PADI-web
reactivity_pw <- calc_reactivity(PATHS_PW)
table(reactivity_pw$reactivity)/nrow(reactivity_pw)

# HealthMap
reactivity_hm <- calc_reactivity(PATHS_HM)
table(reactivity_hm$reactivity)/nrow(reactivity_hm)
```
### Length 

```{r , echo=FALSE}
# PADI-web
length_pw <- PATHS_PW %>% count(id_path)
rbind(table(length_pw$n),table(length_pw$n)/nrow(length_pw))

# HealthMap
length_hm <- PATHS_HM %>% count(id_path)
rbind(table(length_hm$n),table(length_hm$n)/nrow(length_hm))

```

### 2.4. Sources 

#### Sources in PADI-web network
```{r}
SOURCES_PW <- c(PATHS_PW$source_1, PATHS_PW$source_2) %>% unique()
SOURCES_PW <- merge(data.frame(source = SOURCES_PW),
                     select(SOURCES, label, id, type_source), 
                     by.x = 'source', by.y = 'label',all.x=TRUE, all.y=FALSE)                                 
SOURCES_PW %>% filter(!is.na(type_source))  %>% 
  count(type_source) %>% 
  arrange(desc(n)) %>%          
  mutate(prop = prop.table(n))                                

```

#### Sources in HealthMap network
```{r}
SOURCES_HM <- c(PATHS_HM$source_1, PATHS_HM$source_2) %>% unique()
SOURCES_HM <- merge(data.frame(source = SOURCES_HM),
                     select(SOURCES, label, id, type_source), 
                     by.x = 'source', by.y = 'label',all.x=TRUE, all.y=FALSE)                                 
SOURCES_HM %>% 
  count(type_source) %>% 
  arrange(desc(n)) %>%          
  mutate(prop = prop.table(n))                                

```

#### Sources specific to PADI-web network
```{r}
n_unique = SOURCES_PW$source[!SOURCES_PW$source %in% SOURCES_HM$source] %>% length()

rbind(SOURCES_PW %>% filter(!source %in% SOURCES_HM$source) %>%
  count(type_source) %>% 
  arrange(desc(n)) %>%          
  mutate(prop = prop.table(n)),
  c("total", n_unique, 1))

```
#### Sources specific to HealthMap network
```{r}
n_unique = SOURCES_HM$source[!SOURCES_HM$source %in% SOURCES_HM$source] %>% length()

rbind(SOURCES_HM %>% filter(!source %in% SOURCES_PW$source) %>%
  count(type_source) %>% 
  arrange(desc(n)) %>%          
  mutate(prop = prop.table(n)),
  c("total", n_unique, 1))

```


# 3. Qualitative network analysis

## 3.1. Primary sources (emettor)

```{r}

# Late paths
id_path_late_pw <- pull(filter(PATHS_EVENTS_OFF_PW, timeliness >= 0), id_path)
PATHS_LATE_PW <- filter(PATHS_PW, id_path %in% id_path_late_pw)

id_path_late_hm <- pull(filter(PATHS_EVENTS_OFF_HM, timeliness >= 0), id_path)
PATHS_LATE_HM<- filter(PATHS_HM, id_path %in% id_path_late_hm)

primary_late_pw <- calc_primary_source(PATHS_LATE_PW, SOURCES)
primary_late_hm <- calc_primary_source(PATHS_LATE_HM, SOURCES)

# Early paths
id_path_early_pw <- pull(filter(PATHS_EVENTS_OFF_PW, timeliness <0), id_path)
PATHS_EARLY_PW <- filter(PATHS_PW, id_path %in% id_path_early_pw)

id_path_early_hm <- pull(filter(PATHS_EVENTS_OFF_HM, timeliness < 0), id_path)
PATHS_EARLY_HM<- filter(PATHS_HM, id_path %in% id_path_early_hm)

primary_early_pw<- calc_primary_source(PATHS_EARLY_PW, SOURCES) 
primary_early_hm<- calc_primary_source(PATHS_EARLY_HM, SOURCES) 


# Non-official paths
id_events_unoff <- unique(c(pull(filter(EVENTS_PW, type_event == "unofficial"), id_event),
pull(filter(EVENTS_HM, type_event == "unofficial"), id_event)))

id_path_unoff_pw <- unique(pull(filter(PATHS_EVENTS_PW, id_event %in% id_events_unoff), id_path))
PATHS_UNOFF_PW <- filter(PATHS_PW, id_path %in% id_path_unoff_pw)

id_path_unoff_hm <- unique(pull(filter(PATHS_EVENTS_HM, id_event %in% id_events_unoff), id_path))
PATHS_UNOFF_HM <- filter(PATHS_HM, id_path %in% id_path_unoff_hm)


primary_unoff_pw<- calc_primary_source(PATHS_UNOFF_PW, SOURCES) 
primary_unoff_hm<- calc_primary_source(PATHS_UNOFF_HM, SOURCES) 

```

Primary sources aggregated per type

```{r, warning = F}
excluded_sources <- c("local person", "research org.", "private company")
primary_late_pw_aggr <- aggregate_source(primary_late_pw, exclude = excluded_sources)
primary_late_hm_aggr <- aggregate_source(primary_late_hm, exclude = excluded_sources)

primary_early_pw_aggr <- aggregate_source(primary_early_pw, exclude = excluded_sources)
primary_early_hm_aggr <- aggregate_source(primary_early_hm, exclude = excluded_sources)

primary_unoff_pw_aggr <- aggregate_source(primary_unoff_pw)
primary_unoff_hm_aggr <- aggregate_source(primary_unoff_hm)

```


## 3.2. Secondary sources (emettor-receptor)

```{r}

# Late paths

secondary_late_pw <- calc_secondary_source(PATHS_LATE_PW, SOURCES)
secondary_late_hm <- calc_secondary_source(PATHS_LATE_HM, SOURCES)

# Early paths
secondary_early_pw<- calc_secondary_source(PATHS_EARLY_PW, SOURCES) 
secondary_early_hm<- calc_secondary_source(PATHS_EARLY_HM, SOURCES) 


# Non-official paths
secondary_unoff_pw<- calc_secondary_source(PATHS_UNOFF_PW, SOURCES) 
secondary_unoff_hm<- calc_secondary_source(PATHS_UNOFF_HM, SOURCES) 
```

```{r}
secondary_late_pw_aggr <- aggregate_source(secondary_late_pw, exclude = excluded_sources)
secondary_late_hm_aggr <- aggregate_source(secondary_late_hm, exclude = excluded_sources)

secondary_early_pw_aggr <- aggregate_source(secondary_early_pw, exclude = excluded_sources)
secondary_early_hm_aggr <- aggregate_source(secondary_early_hm, exclude = excluded_sources)


secondary_unoff_pw_aggr <- aggregate_source(secondary_unoff_pw)
secondary_unoff_hm_aggr <- aggregate_source(secondary_unoff_hm)

```

## 3.3. Distribution per type of sources

### Distribution of primary and secondary sources in late detection

```{r, fig.width=10,fig.height=6}
df_plot_pw <- convert_source_plot(primary_late_pw_aggr, secondary_late_pw_aggr, table_convert,  exclude = excluded_sources)
df_plot_hm <- convert_source_plot(primary_late_hm_aggr, secondary_late_hm_aggr, table_convert,  exclude = excluded_sources)
```

```{r}
order_source <- c(
  "National vet. auth.",
  "Local vet. auth.",
  "Online news",
  "Local auth.",
  "Laboratory",
  "National auth.",
  "Radio, TV" ,
  "International vet. auth.",
  "Press agency",
  "Social platform"
)

p <-
  create_source_plot(df_plot_pw, min_prop = 0, order_source = order_source)
g <-
  create_source_plot(
    df_plot_hm ,
    min_prop = 0,
    order_source = order_source,
    position_x_axis = "top",
    legend = F
  )

p %>% ggsave(
  filename = "./figures/pw_late_sources.png",
  width = 6,
  height = 4,
  dpi = 300
)
g %>% ggsave(
  filename = "./figures/hm_late_sources.png",
  width = 6,
  height = 4,
  dpi = 300
)

par(mfrow=c(1,2))
p
g
```

### Distribution of primary and secondary sources in early detection


```{r, fig.width=10,fig.height=6}
df_plot_pw <- convert_source_plot(primary_early_pw_aggr, secondary_early_pw_aggr, table_convert,  exclude = c(excluded_sources, "international vet auth."))
df_plot_hm <- convert_source_plot(primary_early_hm_aggr, secondary_early_hm_aggr, table_convert,  exclude = c(excluded_sources, "international vet auth."))
```

```{r}
order_source <- c(   
                   "Local vet. auth.",
                 "National vet. auth.", 
                      "Local auth.",
                           "Online news",
                   "Radio, TV" ,
                   "Press agency", 
                   "Laboratory",
                              "Social platform",
                       "National auth.")

p <- create_source_plot(df_plot_pw, min_prop = 0, order_source = order_source)
g <- create_source_plot(df_plot_hm , min_prop = 0.0, order_source= order_source, position_x_axis = "top", legend = F)


p
g
p %>% ggsave(filename = "./figures/pw_early_sources.png",
               width = 6, height = 4, dpi = 300)

g %>% ggsave(filename = "./figures/hm_early_sources.png",
               width = 6, height = 4, dpi = 300)
```
### Distribution of primary and secondary sources in the detection of non-official events

```{r, fig.width=10,fig.height=6}
bind_rows(bind_cols(primary_unoff_pw_aggr, "type" = "primary"),
          bind_cols(secondary_unoff_pw_aggr, "type" = "secondary"))
bind_rows(bind_cols(primary_unoff_hm_aggr, "type" = "primary"),
          bind_cols(secondary_unoff_hm_aggr, "type" = "secondary"))

```

## 3.4. Distribution per geographical focus

```{r}
primary_early_pw_focus <- merge(primary_early_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_early_pw_focus_aggr <- aggregate_source(primary_early_pw_focus, field = "geographical_focus")

primary_late_pw_focus <- merge(primary_late_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_late_pw_focus_aggr <- aggregate_source(primary_late_pw_focus, field = "geographical_focus")

primary_early_hm_focus <- merge(primary_early_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_early_hm_focus_aggr <- aggregate_source(primary_early_hm_focus, field = "geographical_focus")

primary_late_hm_focus <- merge(primary_late_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_late_hm_focus_aggr <- aggregate_source(primary_late_hm_focus, field = "geographical_focus")

secondary_early_pw_focus <- merge(secondary_early_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_early_pw_focus_aggr <- aggregate_source(secondary_early_pw_focus, field = "geographical_focus")

secondary_late_pw_focus <- merge(secondary_late_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_late_pw_focus_aggr <- aggregate_source(secondary_late_pw_focus, field = "geographical_focus")

secondary_early_hm_focus <- merge(secondary_early_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_early_hm_focus_aggr <- aggregate_source(secondary_early_hm_focus, field = "geographical_focus")

secondary_late_hm_focus <- merge(secondary_late_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_late_hm_focus_aggr <- aggregate_source(secondary_late_hm_focus, field = "geographical_focus")

```

Primary sources:
```{r}
plot_focus_primary <- create_sp_focus_plot(primary_late_pw_focus_aggr, primary_late_hm_focus_aggr,
                                 primary_early_pw_focus_aggr, primary_early_hm_focus_aggr,
                                 column = "geographical_focus", levels = c("local", "national", "international")) 

plot_focus_primary %>% ggsave(filename = "./figures/focus_primary.png",
             width = 8.5, height = 4, dpi = 300)
plot_focus_primary

```

Secondary sources:
```{r}
plot_focus_secondary <- create_sp_focus_plot(secondary_late_pw_focus_aggr, secondary_late_hm_focus_aggr,
                                 secondary_early_pw_focus_aggr, secondary_early_hm_focus_aggr,
                                 column = "geographical_focus", levels = c("local", "national", "international")) 

plot_focus_secondary %>% ggsave(filename = "./figures/focus_secondary.png",
             width = 8.5, height = 4, dpi = 300)
plot_focus_secondary
```

## 3.5. Distribution per specialization

```{r}
primary_early_pw_spec <- merge(primary_early_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_early_pw_spec_aggr <- aggregate_source(primary_early_pw_spec, field = "specialization")

primary_late_pw_spec <- merge(primary_late_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_late_pw_spec_aggr <- aggregate_source(primary_late_pw_spec, field = "specialization")

primary_early_hm_spec <- merge(primary_early_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_early_hm_spec_aggr <- aggregate_source(primary_early_hm_spec, field = "specialization")

primary_late_hm_spec <- merge(primary_late_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_late_hm_spec_aggr <- aggregate_source(primary_late_hm_spec, field = "specialization")


secondary_early_pw_spec <- merge(secondary_early_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_early_pw_spec_aggr <- aggregate_source(secondary_early_pw_spec, field = "specialization")

secondary_late_pw_spec <- merge(secondary_late_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_late_pw_spec_aggr <- aggregate_source(secondary_late_pw_spec, field = "specialization")

secondary_early_hm_spec <- merge(secondary_early_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_early_hm_spec_aggr <- aggregate_source(secondary_early_hm_spec, field = "specialization")

secondary_late_hm_spec <- merge(secondary_late_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_late_hm_spec_aggr <- aggregate_source(secondary_late_hm_spec, field = "specialization")
```


Primary sources:
```{r}
plot_specialization_primary <- create_sp_focus_plot(primary_late_pw_spec_aggr, primary_late_hm_spec_aggr,
                                           primary_early_pw_spec_aggr, primary_early_hm_spec_aggr,
                                   column = "specialization", levels = c("general", "specialized")) 
plot_specialization_primary 
plot_specialization_primary  %>% ggsave(filename = "./figures/specialization.png",
                       width = 8, height = 4, dpi = 300)



```

Secondary sources:
```{r}
plot_specialization_secondary  <- create_sp_focus_plot(secondary_late_pw_spec_aggr, secondary_late_hm_spec_aggr,
                                           secondary_early_pw_spec_aggr, secondary_early_hm_spec_aggr,
                                   column = "specialization", levels = c("general", "specialized")) 
plot_specialization_secondary
plot_specialization_secondary %>% ggsave(filename = "./figures/specialization_secondary.png",
                       width = 8, height = 4, dpi = 300)



```


# 4. Quantitative network analysis



### 3.1. Graph creation
```{r}

PATHS_NETWORK <- create_df_network(bind_rows(PATHS_PW, PATHS_HM), SOURCES)
g <- graph.data.frame(PATHS_NETWORK, directed = TRUE)

PATHS_NETWORK_PW <- create_df_network(PATHS_PW, SOURCES)
g_pw <- graph.data.frame(PATHS_NETWORK_PW, directed = TRUE)

PATHS_NETWORK_HM <- create_df_network(PATHS_HM, SOURCES)
g_hm <- graph.data.frame(PATHS_NETWORK_HM, directed = TRUE)

```
### 3.2. Graph characteristics

#### Nodes
```{r}
vcount(g_pw)
vcount(g_hm)
```

#### Vertices
```{r}
ecount(g_pw)
ecount(g_hm)
```

#### Paths
```{r}
length(unique(PATHS_PW$id_path))
length(unique(PATHS_HM$id_path))
```

### 3.3. Nodes characteristics   

```{r, warning=F}
nodes <- df_nodes_importance(g, node_remove = NULL)
nodes_type <- merge(nodes, SOURCES, by.x = "source", by.y = "label", all.x = T) %>% filter(!is.na(type_source))
plot_in <- boxplot_sources(nodes_type)
plot_out <- boxplot_sources(nodes_type, type = "out")
plot_all <- boxplot_sources(nodes_type, type = "all")
plot_in %>% ggsave(filename = "./figures/in_degree.png",
                    width = 8, height = 4, dpi = 300)
plot_out %>% ggsave(filename = "./figures/out_degree.png",
                    width = 8, height = 4, dpi = 300)
plot_all %>% ggsave(filename = "./figures/all_degree.png",
                    width = 8, height = 4, dpi = 300)

```

```{r, warning=F}
par(mfrow=c(1,3)) 
plot_in 
plot_out 
plot_all
```

Most important sources in terms of degree or centrality values


Higest in-degree
```{r}
nodes %>% slice_max(in_degree, n = 5)
```

Higest out-degree
```{r}
nodes %>% slice_max(out_degree, n = 5)
```

Higest all-degree
```{r}
nodes %>% slice_max(all_degree, n = 5)
```

### 4. Network visualisation 
```{r}
PATHS_PW_DIAGRAM <- PATHS_PW  %>% filter(!source_1=="local person" & !source_2=="local person") %>% select(source_2, source_1 ) %>% group_by(source_2, source_1 ) %>% count()
names(PATHS_PW_DIAGRAM) <- c("Source", "Target", "weight")
write.table(PATHS_PW_DIAGRAM, "./dataverse/paths_pw_diagram.csv", sep=";", row.names = F, col.names = T)


PATHS_HM_DIAGRAM <- PATHS_HM  %>% select(source_2, source_1 ) %>% group_by(source_2, source_1 ) %>% count()
names(PATHS_HM_DIAGRAM) <- c("Source", "Target", "weight")
write.table(PATHS_HM_DIAGRAM, "./dataverse/paths_hm_diagram.csv", sep=";", row.names = F, col.names = T)

```

```{r}
padiweb.output.filepath = "./figures/chord_diagram_padiweb.png"
padiweb.output.nodes.legend = "./figures/chord_diagram_padiweb_nodes_legend.csv"

healthmap.output.filepath = "./figures/chord_diagram_healthmap.png"
healthmap.output.nodes.legend = "./figures/chord_diagram_healthmap_nodes_legend.csv"

padiweb.links.filepath = "./dataverse/links_padiweb.csv"
healthmap.links.filepath = "./dataverse/links_healthmap.csv"

padiweb.edges.csv.filepath = "./dataverse/paths_pw_diagram.csv"
padiweb.info.csv.filepath = "./dataverse/sources.csv"
healthmap.edges.csv.filepath = "./dataverse/paths_hm_diagram.csv"
healthmap.info.csv.filepath = "./dataverse/sources.csv"

groups = retrieve.group.names(padiweb.info.csv.filepath)
groups.adjusted = sapply(groups, function(x){ parts = split.text.into.two.parts(x); return(paste0(parts[1],"\n",parts[2]))}) # make 2-line group names for the sake of readibility in the plot, the same as in the function 'prepare.input.for.chord.diagram()' >> for color assignment, we need to reproduce these text adjustments here, as well
colors.for.groups = create.colors.group.association(groups.adjusted)

new_levels <- c("international vet auth.", "EBS tool",
                "social platform", "press agency",
                "online news source", "research org.", 
                "radio, TV",  "national vet auth.",
                "laboratory", "local off. auth.", 
                "local vet auth.", "private company",                   
                "national off. auth.")

# Fill colour
colors.for.group <- c(brewer.pal(8,'Dark2'), brewer.pal(name="Paired", n = length(new_levels)-8))
names(colors.for.group) <- new_levels
```

Padiweb

```{r}
prepare.input.for.chord.diagram(padiweb.edges.csv.filepath, padiweb.info.csv.filepath, padiweb.links.filepath) # it creates 'padiweb.links.filepath'
small.gap = 0.5
big.gap = 1.8 
create.chord.diagram(padiweb.links.filepath, padiweb.output.filepath, padiweb.output.nodes.legend,
                     colors.for.group,  small.gap, 
                     big.gap, include_legend = F)
```

Healthmap

```{r}
prepare.input.for.chord.diagram(healthmap.edges.csv.filepath, healthmap.info.csv.filepath, healthmap.links.filepath) # it creates 'healthmap.links.filepath'
small.gap = 0.4
create.chord.diagram(healthmap.links.filepath, healthmap.output.filepath, healthmap.output.nodes.legend, colors.for.group, small.gap, big.gap, include_legend = T)
```

