---
title: "Event-based surveillance - network analysis (AI)"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=F}
suppressMessages(library(readxl))
library(reshape2)
library(dplyr)
library(ggplot2)
library(igraph)
library(stringr)
library(ggpubr)
library(readr)
library(RColorBrewer)
setwd("C:/GitHub/network_ebs")
source("./R/sna_functions.R")
source("./R/create_chord_diagrams.R")
```


# 1. Data Import

Data can be imported from:

```{r, message=FALSE, warning=F}

# Reports from PADI-web
REPORTS_PW <- read_delim("./data/padiweb_reports.csv", show_col_types = FALSE) %>% mutate(publication_date = as.Date(publication_date, format="%Y-%m-%d"))

# Reports from HealthMap
REPORTS_HM <- read_delim("./data/healthmap_reports.csv", show_col_types = FALSE) %>% mutate(publication_date = as.Date(publication_date, format="%Y-%m-%d"))

# Events empres-i database
OIE_EVENTS  <- open_official_AI("./data/AI_official.csv", date_format =  "%Y-%m-%d")
OIE_EVENTS <- filter(OIE_EVENTS, reportingdate >= "2018-07-01") %>% filter(reportingdate <= "2019-06-30")

# Manually recorded non-official events
UNOFFICIAL_EVENTS <- read_delim("./data/unofficial_events.csv", 
                                delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)

# Sources
SOURCES <- read_delim(file="./data/sources.csv", delim=";", show_col_types = F) %>% mutate(label = str_trim(label))

# Tables for PADI-web network
EVENTS_PW <- read_delim("./data/padiweb_events.csv", 
                        delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)
PATHS_PW <- read_delim("./data/padiweb_paths.csv", delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)
PATHS_PW$source_1 <- str_trim(PATHS_PW$source_1)
PATHS_PW$source_2 <- str_trim(PATHS_PW$source_2)
PATHS_EVENTS_PW <- read_delim("./data/padiweb_paths_events.csv", delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)

# Tables for HealthMap network
EVENTS_HM <- read_delim("./data/healthmap_events.csv", 
                        delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)
PATHS_HM <- read_delim("./data/healthmap_paths.csv", delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)
PATHS_HM$source_1 <- str_trim(PATHS_HM$source_1)
PATHS_HM$source_2 <- str_trim(PATHS_HM$source_2)
PATHS_HM$id_from_source_1 <- as.character(PATHS_HM$id_from_source_1)
PATHS_EVENTS_HM <- read_delim("./data/healthmap_paths_events.csv", delim  = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = F)

```


# 2. Descriptive statistics

## 2.1. Events

### Official events from Empres-i
```{r, message=FALSE, warning=F}
rbind(OIE_EVENTS %>% count(type_host) %>% 
        mutate(prop = prop.table(n)),
      c("total", nrow(OIE_EVENTS), 1))
```


### Events disseminated through the networks
```{r}
# PADI-web
EVENTS_PW <- EVENTS_PW %>% filter(id_event %in% PATHS_EVENTS_PW$id_event) 
rbind(EVENTS_PW %>% select(id_event, type_event, host) %>% distinct() %>% count(type_event) %>% 
        mutate(prop = round(prop.table(n),3)),
      c("total", nrow(EVENTS_PW), 1))

# HealthMap
EVENTS_HM <- EVENTS_HM %>% filter(id_event %in% PATHS_EVENTS_HM$id_event) 
rbind(EVENTS_HM %>% select(id_event, type_event, host) %>% distinct() %>% count(type_event) %>% 
        mutate(prop = round(prop.table(n),3)),
      c("total", nrow(EVENTS_HM), 1))
```

### Overlap of detected events
```{r}
# PADI-web
EVENTS_PW %>% filter(!id_event %in% EVENTS_HM$id_event) %>% select(id_event, type_event, type_host) %>% group_by(type_event) %>% distinct() %>% count(type_host)

# HealthMap
EVENTS_HM %>% filter(!id_event %in% EVENTS_PW$id_event) %>% select(id_event, type_event, type_host) %>% group_by(type_event) %>% distinct() %>% count(type_host)

```

## 2.2 Host

```{r}
# PADI-web
EVENTS_PW %>% select(id_event, type_event, type_host) %>% group_by(type_event) %>% distinct() %>% count(type_host)%>% 
  mutate(prop = round(prop.table(n),3))

# HealthMap
EVENTS_HM %>% select(id_event, type_event, type_host) %>% group_by(type_event) %>% distinct() %>% count(type_host)%>% 
  mutate(prop = round(prop.table(n),3))
```

## 2.3 Reports

### Types of reports
```{r, message=FALSE}
# PADI-web
rbind(REPORTS_PW %>% count(type) %>% 
        mutate(prop = prop.table(n)) %>% arrange(desc(prop)),
      c("total", nrow(REPORTS_PW), 1))

# HealthMap
rbind(REPORTS_HM %>% count(type) %>% 
        mutate(prop = prop.table(n)) %>% arrange(desc(prop)),
      c("total", nrow(REPORTS_HM), 1))
```


### Events of reports
```{r}
PATHS_EVENTS_PW  <- merge(PATHS_EVENTS_PW, EVENTS_PW)
REPORTS_EVENTS_PW <- merge(PATHS_EVENTS_PW, select(PATHS_PW, id_path, id_from_source_1))  %>% distinct()

event_by_report_pw <- aggregate(select(REPORTS_EVENTS_PW, id_from_source_1),
                                by = list(id_article= REPORTS_EVENTS_PW$id_from_source_1), 
                                FUN = length)


PATHS_EVENTS_HM  <- merge(PATHS_EVENTS_HM, EVENTS_HM)
REPORTS_EVENTS_HM <-merge(PATHS_EVENTS_HM, select(PATHS_HM, id_path, id_from_source_1))%>% distinct()
event_by_report_hm <- aggregate(select(REPORTS_EVENTS_HM, id_from_source_1),
                                by = list(id_article = REPORTS_EVENTS_HM$id_from_source_1), 
                                FUN = length)

summary(event_by_report_pw$id_from_source_1)
summary(event_by_report_hm$id_from_source_1)

```

## 2.4. Timeliness


```{r}
# Official events
id_events_off <- pull(filter(EVENTS_PW, type_event == "official"), id_event)

# Paths transmitting official events
id_paths_off <- unique(pull(filter(PATHS_EVENTS_PW, id_event %in% id_events_off), id_path))

PATHS_EVENTS_OFF_PW <- filter(PATHS_EVENTS_PW, id_path %in% id_paths_off) %>% filter(id_event %in% id_events_off)

PATHS_EVENTS_OFF_PW <- merge(PATHS_EVENTS_OFF_PW, EVENTS_PW, by.all = "id_event", all.y =F) 

PATHS_EVENTS_OFF_PW <- merge(PATHS_EVENTS_OFF_PW, select(PATHS_PW, id_path, source_1, date_edge), by.all = "id_path")

PATHS_EVENTS_OFF_PW <- PATHS_EVENTS_OFF_PW %>% filter(source_1 %in% c("PADI-web"))

PATHS_EVENTS_OFF_PW$timeliness <- as.numeric(difftime(PATHS_EVENTS_OFF_PW$date_edge, 
                                                      PATHS_EVENTS_OFF_PW$reportingdate,
                                                      units="days"))


# For each event, detect the first report which detected it
df_timeliness_pw <- PATHS_EVENTS_OFF_PW %>% select(id_event, type_host, timeliness) %>%
  group_by(id_event, type_host) %>% 
  filter(timeliness == min(timeliness)) %>% unique()

summary(df_timeliness_pw$timeliness)

# Early detection
length(which(df_timeliness_pw$timeliness<0))

# Late detection
length(which(df_timeliness_pw$timeliness>=0))

```
```{r}
host_levels <-  c("domestic", "wild", "environmental", "unspecified")
df_timeliness_pw <- df_timeliness_pw %>% mutate(Host = factor(type_host, levels = host_levels)) %>% group_by(timeliness, Host) %>% count() %>% mutate(n = n*100/nrow(df_timeliness_pw))

nrow(filter(PATHS_EVENTS_OFF_PW, type_host == "wild") %>% filter(timeliness<0))
nrow(filter(PATHS_EVENTS_OFF_PW, type_host == "domestic") %>% filter(timeliness<0))
```

```{r}
# Official events
id_events_off <- pull(filter(EVENTS_HM, type_event == "official"), id_event)

# Paths transmitting official events
id_paths_off <- unique(pull(filter(PATHS_EVENTS_HM, id_event %in% id_events_off), id_path))

PATHS_EVENTS_OFF_HM <- filter(PATHS_EVENTS_HM, id_path %in% id_paths_off) %>% filter(id_event %in% id_events_off)

PATHS_EVENTS_OFF_HM <- merge(PATHS_EVENTS_OFF_HM, EVENTS_HM, by.all = "id_event", all.y =F) 

PATHS_EVENTS_OFF_HM <- merge(PATHS_EVENTS_OFF_HM, select(PATHS_HM, id_path, source_1, date_edge), by.all = "id_path")

PATHS_EVENTS_OFF_HM <- PATHS_EVENTS_OFF_HM %>% filter(source_1 %in% c("HealthMap"))

PATHS_EVENTS_OFF_HM$timeliness <- as.numeric(difftime(PATHS_EVENTS_OFF_HM$date_edge, 
                                                      PATHS_EVENTS_OFF_HM$reportingdate,
                                                      units="days"))

# For each event, detect the first report which detected it
df_timeliness_hm <- PATHS_EVENTS_OFF_HM %>% select(id_event, type_host, timeliness) %>%
  group_by(id_event, type_host) %>% 
  filter(timeliness == min(timeliness)) %>% unique()

df_timeliness_hm <- df_timeliness_hm %>% mutate(Host = factor(type_host, levels = host_levels))


df_timeliness_hm <- df_timeliness_hm %>% select(-id_event) %>% group_by(timeliness, Host) %>% count() %>% mutate(n = n*100/nrow(df_timeliness_hm))


summary(PATHS_EVENTS_OFF_HM$timeliness)

# Early detection
length(which(PATHS_EVENTS_OFF_HM$timeliness<0))

# Late detection
length(which(PATHS_EVENTS_OFF_HM$timeliness>=0))
```
```{r}
nrow(filter(PATHS_EVENTS_OFF_HM, type_host == "wild") %>% filter(timeliness<0))
nrow(filter(PATHS_EVENTS_OFF_HM, type_host == "domestic") %>% filter(timeliness<0))
```

```{r, warning=F}
timeliness_pw <- create_timeliness_plot(df_timeliness_pw, min_timeliness = -30, max_timeliness = 30, size_font = 14, label = "A)")
timeliness_hm <-create_timeliness_plot(df_timeliness_hm, min_timeliness = -30, max_timeliness = 30, size_font = 14, label = "B)")

```

```{r , warning=FALSE}
ggarrange(timeliness_pw, timeliness_hm, nrow=2, common.legend = TRUE, legend="bottom") %>% ggsave(
  filename = "./figures/Fig2.png",
  width = 8,
  height = 8,
  dpi = 300
)

```

## 2.5. Paths metrics

### Reactivity                        

```{r , echo=FALSE}
# PADI-web
reactivity_pw <- calc_reactivity(PATHS_PW)
table(reactivity_pw$reactivity)/nrow(reactivity_pw)

# HealthMap
reactivity_hm <- calc_reactivity(PATHS_HM)
table(reactivity_hm$reactivity)/nrow(reactivity_hm)
```
### Length 

```{r , echo=FALSE}
# PADI-web
length_pw <- PATHS_PW %>% count(id_path)
rbind(table(length_pw$n),table(length_pw$n)/nrow(length_pw))

# HealthMap
length_hm <- PATHS_HM %>% count(id_path)
rbind(table(length_hm$n),table(length_hm$n)/nrow(length_hm))

```

### 2.4. Sources 

#### Sources in PADI-web network
```{r}
SOURCES_PW <- c(PATHS_PW$source_1, PATHS_PW$source_2) %>% unique()
SOURCES_PW <- merge(data.frame(source = SOURCES_PW),
                    select(SOURCES, label, id, type_source), 
                    by.x = 'source', by.y = 'label',all.x=TRUE, all.y=FALSE)                                 
SOURCES_PW %>% filter(!is.na(type_source))  %>% 
  count(type_source) %>% 
  arrange(desc(n)) %>%          
  mutate(prop = prop.table(n))                                

```

#### Sources in HealthMap network
```{r}
SOURCES_HM <- c(PATHS_HM$source_1, PATHS_HM$source_2) %>% unique()
SOURCES_HM <- merge(data.frame(source = SOURCES_HM),
                    select(SOURCES, label, id, type_source), 
                    by.x = 'source', by.y = 'label',all.x=TRUE, all.y=FALSE)                                 
SOURCES_HM %>% 
  count(type_source) %>% 
  arrange(desc(n)) %>%          
  mutate(prop = prop.table(n))                                

```

#### Sources specific to PADI-web network
```{r}
n_unique = SOURCES_PW$source[!SOURCES_PW$source %in% SOURCES_HM$source] %>% length()

padiweb_sources <- merge(SOURCES_PW%>% count(type_source),
  SOURCES_PW %>% filter(!source %in% SOURCES_HM$source) %>%
        count(type_source) %>% 
        rename(n_unique = n) %>% 
        arrange(desc(n_unique)) %>%          
        mutate(prop_unique = prop.table(n_unique)),
  all.x = T)
write.table(padiweb_sources, "./tables/df_padiweb_sources.csv", sep = ";", row.names = F, col.names = T)

```
#### Sources specific to HealthMap network
```{r}
n_unique <- SOURCES_HM$source[!SOURCES_HM$source %in% SOURCES_PW$source] %>% length()

healthmap_sources <- merge(SOURCES_HM%>% count(type_source),
  SOURCES_HM %>% filter(!source %in% SOURCES_PW$source) %>%
        count(type_source) %>% 
        rename(n_unique = n) %>% 
        arrange(desc(n_unique)) %>%          
        mutate(prop_unique = prop.table(n_unique)),
  all.x = T)
write.table(healthmap_sources, "./tables/df_hm_sources.csv", sep = ";", row.names = F, col.names = T)

```


# 3. Qualitative network analysis

## 3.1. Primary sources (emettor)

```{r}

# Late paths
id_path_late_pw <- pull(filter(PATHS_EVENTS_OFF_PW, timeliness >= 0), id_path)
PATHS_LATE_PW <- filter(PATHS_PW, id_path %in% id_path_late_pw)

id_path_late_hm <- pull(filter(PATHS_EVENTS_OFF_HM, timeliness >= 0), id_path)
PATHS_LATE_HM<- filter(PATHS_HM, id_path %in% id_path_late_hm)

primary_late_pw <- calc_primary_source(PATHS_LATE_PW, SOURCES)
primary_late_hm <- calc_primary_source(PATHS_LATE_HM, SOURCES)

# Early paths
id_path_early_pw <- pull(filter(PATHS_EVENTS_OFF_PW, timeliness <0), id_path)
PATHS_EARLY_PW <- filter(PATHS_PW, id_path %in% id_path_early_pw)

id_path_early_hm <- pull(filter(PATHS_EVENTS_OFF_HM, timeliness < 0), id_path)
PATHS_EARLY_HM<- filter(PATHS_HM, id_path %in% id_path_early_hm)

primary_early_pw<- calc_primary_source(PATHS_EARLY_PW, SOURCES) 
primary_early_hm<- calc_primary_source(PATHS_EARLY_HM, SOURCES) 

```

## 3.2. Secondary sources (emettor-receptor)

```{r}

# Late paths

secondary_late_pw <- calc_secondary_source(PATHS_LATE_PW, SOURCES)
secondary_late_hm <- calc_secondary_source(PATHS_LATE_HM, SOURCES)

# Early paths
secondary_early_pw<- calc_secondary_source(PATHS_EARLY_PW, SOURCES) 
secondary_early_hm<- calc_secondary_source(PATHS_EARLY_HM, SOURCES) 

```


## 3.3. Distribution per type of sources

### Distribution of primary and secondary sources in late detection

```{r}
# 
table_convert <- data.frame(
  "type_source" = c(
    "press agency",
    "online news source",
    "international vet auth.",
    "national vet auth.",
    "radio, TV",
    "private company",
    "research org.",
    "laboratory",
    "local off. auth.",
    "national off. auth.",
    "social platform",
    "local vet auth."
  ),
  "name_style" = c(
    "Press agency",
    "Online news",
    "International vet. auth.",
    "National vet. auth.",
    "Radio, TV" ,
    "Private company",
    "Research org.",
    "Laboratory",
    "Local auth.",
    "National auth.",
    "Social platform",
    "Local vet. auth."
  )
)

```

```{r}
df_plot_late_pw <- convert_source_plot(primary_late_pw, secondary_late_pw, field = "type_source", table_convert)

df_plot_late_hm <- convert_source_plot(primary_late_hm, secondary_late_hm, field = "type_source", table_convert)

# private company are at 0 for both role of source in both tools, we remove them 
df_plot_late_pw <- df_plot_late_pw %>% filter(type_source != "private company")
df_plot_late_hm <- df_plot_late_hm %>% filter(type_source != "private company")
```

```{r}
order_source <- c(
  "National vet. auth.",
  "Local vet. auth.",
  "Online news",
  "Local auth.",
  "Laboratory",
  "National auth.",
  "Radio, TV" ,
  "Research org.",
  "Press agency",
  "International vet. auth.",
  "Social platform"
)

plot_late_pw <-
  create_source_plot(df_plot_late_pw, min_prop = 0, order_source = order_source,
                     label = "A)",   x_label=11.3, y_label=-0.65)
plot_late_hm <-
  create_source_plot(
    df_plot_late_hm ,
    order_source = order_source,
    position_x_axis = "top",
    y_axis = F,
    label = "B)",   x_label=11.3, y_label=-0.65,
    legend = F
  )

ggarrange(plot_late_pw, plot_late_hm, nrow=1, common.legend = TRUE) %>% ggsave(
  filename = "./figures/Fig5.png",
  width = 12,
  height = 5,
  dpi = 300
)
```

### Distribution of primary and secondary sources in early detection


```{r, fig.width=10,fig.height=6}
df_plot_early_pw <- convert_source_plot(primary_early_pw, secondary_early_pw, field = "type_source", table_convert)
df_plot_early_hm <- convert_source_plot(primary_early_hm, secondary_early_hm, field = "type_source", table_convert)

# private company are at 0 for both role of source in both tools, we remove them 
df_plot_early_pw <- df_plot_early_pw %>% filter(!type_source %in% c("international vet auth.", "research org.", "local person"))
df_plot_early_hm <- df_plot_early_hm  %>% filter(!type_source %in% c("international vet auth.", "research org."))


```

```{r}
order_source <- c(
  "Local vet. auth.",
  "National vet. auth.",
  "Local auth.",
  "Online news",
  "Radio, TV" ,
  "Private company",
  "Laboratory",
  "National auth.",
  "Press agency",
  "Social platform"
)

plot_early_pw <-
  create_source_plot(df_plot_early_pw, min_prop = 0, order_source = order_source,
                     label = "A)",   x_label=10.3, y_label=-0.5)
plot_early_hm <-
  create_source_plot(
    df_plot_early_hm ,
    order_source = order_source,
    position_x_axis = "top",
    y_axis = F,
    label = "B)",   x_label=10.3, y_label=-0.5,
    legend = F
  )

ggarrange(plot_early_pw, plot_early_hm, nrow=1, common.legend = TRUE) %>% ggsave(
  filename = "./figures/Fig6.png",
  width = 12,
  height = 5,
  dpi = 300
)

```

Export reference tables for the graphics
```{r}
write.table(df_plot_late_pw, "./tables/df_Fig5a.csv", sep = ";", row.names = F, col.names = T)
write.table(df_plot_late_hm, "./tables/df_Fig5b.csv", sep = ";", row.names = F, col.names = T)

write.table(df_plot_early_pw, "./tables/df_Fig6a.csv", sep = ";", row.names = F, col.names = T)
write.table(df_plot_early_hm, "./tables/df_Fig6b.csv", sep = ";", row.names = F, col.names = T)

```

### Distribution of primary and secondary sources in the detection of non-official events

```{r}
id_events_unoff <- unique(c(pull(filter(EVENTS_PW, type_event == "unofficial"), id_event),
                            pull(filter(EVENTS_HM, type_event == "unofficial"), id_event)))

id_path_unoff_pw <- unique(pull(filter(PATHS_EVENTS_PW, id_event %in% id_events_unoff), id_path))
PATHS_UNOFF_PW <- filter(PATHS_PW, id_path %in% id_path_unoff_pw)

id_path_unoff_hm <- unique(pull(filter(PATHS_EVENTS_HM, id_event %in% id_events_unoff), id_path))
PATHS_UNOFF_HM <- filter(PATHS_HM, id_path %in% id_path_unoff_hm)



# Non-official paths
secondary_unoff_pw<- calc_secondary_source(PATHS_UNOFF_PW, SOURCES) 
secondary_unoff_hm<- calc_secondary_source(PATHS_UNOFF_HM, SOURCES) 

primary_unoff_pw<- calc_primary_source(PATHS_UNOFF_PW, SOURCES) 
primary_unoff_hm<- calc_primary_source(PATHS_UNOFF_HM, SOURCES) 
secondary_unoff_pw_aggr <- aggregate_source(secondary_unoff_pw)
secondary_unoff_hm_aggr <- aggregate_source(secondary_unoff_hm)
primary_unoff_pw_aggr <- aggregate_source(primary_unoff_pw)
primary_unoff_hm_aggr <- aggregate_source(primary_unoff_hm)

bind_rows(bind_cols(primary_unoff_pw_aggr, "type" = "primary"),
          bind_cols(secondary_unoff_pw_aggr, "type" = "secondary"))
bind_rows(bind_cols(primary_unoff_hm_aggr, "type" = "primary"),
          bind_cols(secondary_unoff_hm_aggr, "type" = "secondary"))

```

## 3.4. Distribution per geographical focus

```{r}
primary_early_pw_focus <- merge(primary_early_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_early_pw_focus_aggr <- aggregate_source(primary_early_pw_focus, field = "geographical_focus")

primary_late_pw_focus <- merge(primary_late_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_late_pw_focus_aggr <- aggregate_source(primary_late_pw_focus, field = "geographical_focus")

primary_early_hm_focus <- merge(primary_early_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_early_hm_focus_aggr <- aggregate_source(primary_early_hm_focus, field = "geographical_focus")

primary_late_hm_focus <- merge(primary_late_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_late_hm_focus_aggr <- aggregate_source(primary_late_hm_focus, field = "geographical_focus")

secondary_early_pw_focus <- merge(secondary_early_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_early_pw_focus_aggr <- aggregate_source(secondary_early_pw_focus, field = "geographical_focus")

secondary_late_pw_focus <- merge(secondary_late_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_late_pw_focus_aggr <- aggregate_source(secondary_late_pw_focus, field = "geographical_focus")

secondary_early_hm_focus <- merge(secondary_early_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_early_hm_focus_aggr <- aggregate_source(secondary_early_hm_focus, field = "geographical_focus")

secondary_late_hm_focus <- merge(secondary_late_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_late_hm_focus_aggr <- aggregate_source(secondary_late_hm_focus, field = "geographical_focus")

```

Primary sources:
```{r}
plot_focus_primary <- create_sp_focus_plot(primary_late_pw_focus_aggr, primary_late_hm_focus_aggr,
                                           primary_early_pw_focus_aggr, primary_early_hm_focus_aggr,
                                           column = "geographical_focus", levels = c("local", "national", "international")) 

plot_focus_primary

```

Secondary sources:
```{r}
plot_focus_secondary <- create_sp_focus_plot(secondary_late_pw_focus_aggr, secondary_late_hm_focus_aggr,
                                             secondary_early_pw_focus_aggr, secondary_early_hm_focus_aggr,
                                             column = "geographical_focus", levels = c("local", "national", "international"),
           common_label = "Secondary sources") 

plot_focus_secondary
```
```{r}

ggarrange(plot_focus_primary, plot_focus_secondary, nrow=2, common.legend = TRUE, legend="right") %>% ggsave(
  filename = "./figures/Fig7.png",
  width = 10,
  height = 8,
  dpi = 300
)
```


## 3.5. Distribution per specialization

```{r}
primary_early_pw_spec <- merge(primary_early_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_early_pw_spec_aggr <- aggregate_source(primary_early_pw_spec, field = "specialization")

primary_late_pw_spec <- merge(primary_late_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_late_pw_spec_aggr <- aggregate_source(primary_late_pw_spec, field = "specialization")

primary_early_hm_spec <- merge(primary_early_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_early_hm_spec_aggr <- aggregate_source(primary_early_hm_spec, field = "specialization")

primary_late_hm_spec <- merge(primary_late_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
primary_late_hm_spec_aggr <- aggregate_source(primary_late_hm_spec, field = "specialization")


secondary_early_pw_spec <- merge(secondary_early_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_early_pw_spec_aggr <- aggregate_source(secondary_early_pw_spec, field = "specialization")

secondary_late_pw_spec <- merge(secondary_late_pw, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_late_pw_spec_aggr <- aggregate_source(secondary_late_pw_spec, field = "specialization")

secondary_early_hm_spec <- merge(secondary_early_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_early_hm_spec_aggr <- aggregate_source(secondary_early_hm_spec, field = "specialization")

secondary_late_hm_spec <- merge(secondary_late_hm, select(SOURCES, label, geographical_focus, specialization), by.x = "source", by.y = "label", all.x = T)
secondary_late_hm_spec_aggr <- aggregate_source(secondary_late_hm_spec, field = "specialization")
```


Primary sources:
```{r}
plot_specialization_primary <- create_sp_focus_plot(primary_late_pw_spec_aggr, primary_late_hm_spec_aggr,
                                                    primary_early_pw_spec_aggr, primary_early_hm_spec_aggr,
                                                    column = "specialization", levels = c("general", "specialized")) 
plot_specialization_primary 

```

Secondary sources:
```{r}
plot_specialization_secondary  <- create_sp_focus_plot(secondary_late_pw_spec_aggr, secondary_late_hm_spec_aggr,
                                                       secondary_early_pw_spec_aggr, secondary_early_hm_spec_aggr,
                                                       column = "specialization", levels = c("general", "specialized")) 
plot_specialization_secondary

```
```{r}

ggarrange(plot_specialization_primary, plot_specialization_secondary, nrow=2, common.legend = TRUE, legend="right") %>% ggsave(
  filename = "./figures/FigSupp1.png",
  width = 10,
  height = 8,
  dpi = 300
)
```

# 4. Quantitative network analysis



### 3.1. Graph creation
```{r}

PATHS_NETWORK <- create_df_network(bind_rows(PATHS_PW, PATHS_HM), SOURCES)
g <- graph.data.frame(PATHS_NETWORK, directed = TRUE)

PATHS_NETWORK_PW <- create_df_network(PATHS_PW, SOURCES)
g_pw <- graph.data.frame(PATHS_NETWORK_PW, directed = TRUE)

PATHS_NETWORK_HM <- create_df_network(PATHS_HM, SOURCES)
g_hm <- graph.data.frame(PATHS_NETWORK_HM, directed = TRUE)

```
### 3.2. Graph characteristics

#### Nodes
```{r}
vcount(g_pw)
vcount(g_hm)
```

#### Vertices
```{r}
ecount(g_pw)
ecount(g_hm)
```

#### Paths
```{r}
length(unique(PATHS_PW$id_path))
length(unique(PATHS_HM$id_path))
```

### 3.3. Nodes characteristics   

```{r, warning=F}
nodes <- df_nodes_importance(g, node_remove = NULL)
nodes_type <-
  merge(nodes,
        SOURCES,
        by.x = "source",
        by.y = "label",
        all.x = T) %>% filter(!is.na(type_source))
plot_in <- boxplot_sources(nodes_type)
plot_out <- boxplot_sources(nodes_type, type = "out")
plot_all <- boxplot_sources(nodes_type, type = "all")

ggarrange(
  plot_in,
  plot_out,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
) %>% ggsave(
  filename = "./figures/Fig4.png",
  width = 8,
  height = 7,
  dpi = 300
)
```

```{r, warning=F}
par(mfrow=c(1,3)) 
plot_in 
plot_out 
plot_all
```

Most important sources in terms of degree or centrality values


Higest in-degree
```{r}
nodes %>% slice_max(in_degree, n = 5)
```

Higest out-degree
```{r}
nodes %>% slice_max(out_degree, n = 5)
```

Higest all-degree
```{r}
nodes %>% slice_max(all_degree, n = 5)
```



### 4. Network visualisation 
```{r}
PATHS_PW_DIAGRAM <-
  PATHS_PW  %>% select(source_2, source_1) %>% group_by(source_2, source_1) %>% count()
names(PATHS_PW_DIAGRAM) <- c("Source", "Target", "weight")
write.table(
  PATHS_PW_DIAGRAM,
  "./tables/paths_pw_diagram.csv",
  sep = ";",
  row.names = F,
  col.names = T
)


PATHS_HM_DIAGRAM <-
  PATHS_HM  %>% select(source_2, source_1) %>% group_by(source_2, source_1) %>% count()
names(PATHS_HM_DIAGRAM) <- c("Source", "Target", "weight")
write.table(
  PATHS_HM_DIAGRAM,
  "./tables/paths_hm_diagram.csv",
  sep = ";",
  row.names = F,
  col.names = T
)

```

```{r}
padiweb.output.filepath = "./figures/Fig3a.png"
padiweb.output.nodes.legend = "./tables/Table_Fig3a.csv"
padiweb.output.df.diagram = "./tables/Table2_Fig3a.csv"

healthmap.output.filepath = "./figures/Fig3b.png"
healthmap.output.nodes.legend = "./tables/Table_Fig3b.csv"
healthmap.output.df.diagram = "./tables/Table2_Fig3b.csv"

padiweb.links.filepath = "./tables/links_padiweb.csv"
healthmap.links.filepath = "./tables/links_healthmap.csv"

padiweb.edges.csv.filepath = "./tables/paths_pw_diagram.csv"
padiweb.info.csv.filepath = "./data/sources.csv"
healthmap.edges.csv.filepath = "./tables/paths_hm_diagram.csv"
healthmap.info.csv.filepath = "./data/sources.csv"

groups = retrieve.group.names(padiweb.info.csv.filepath)
groups.adjusted = sapply(groups, function(x){ parts = split.text.into.two.parts(x); return(paste0(parts[1],"\n",parts[2]))}) # make 2-line group names for the sake of readibility in the plot, the same as in the function 'prepare.input.for.chord.diagram()' >> for color assignment, we need to reproduce these text adjustments here, as well
colors.for.groups = create.colors.group.association(groups.adjusted)

new_levels <- c("international vet auth.", "EBS tool",
                "social platform", "press agency",
                "online news source", "research org.", 
                "radio, TV",  "national vet auth.",
                "laboratory", "local off. auth.", 
                "local vet auth.", "private company",                   
                "national off. auth.")

# Fill colour
colors.for.group <- c(brewer.pal(8,'Dark2'), brewer.pal(name="Paired", n = length(new_levels)-8))
names(colors.for.group) <- new_levels
```

Padiweb

```{r}
prepare.input.for.chord.diagram(padiweb.edges.csv.filepath, padiweb.info.csv.filepath, padiweb.links.filepath) # it creates 'padiweb.links.filepath'
small.gap = 0.5
big.gap = 1.8 
create.chord.diagram(padiweb.links.filepath, padiweb.output.filepath, padiweb.output.nodes.legend,
                     healthmap.output.df.diagram,
                     colors.for.group,  small.gap, 
                     big.gap, include_legend = F, label = "A)")
```

Healthmap

```{r}
prepare.input.for.chord.diagram(healthmap.edges.csv.filepath, healthmap.info.csv.filepath, healthmap.links.filepath) # it creates 'healthmap.links.filepath'
small.gap = 0.4
big.gap = 1.8 
create.chord.diagram(healthmap.links.filepath, healthmap.output.filepath, healthmap.output.nodes.legend, healthmap.output.df.diagram, colors.for.group, small.gap, big.gap, include_legend = T, label = "B)")
```

